Estimating Carrying Capacity for Chinook Parr using Quantile Random Forest Models

### 2020-06-09 analysis/paper/paper.Rmd
updated captions and cross references for photo figures
changed the name of habitat metrics for plotting. Added reference about master sample points.

```{r load-data}
fish_hab = read_rds("../data/derived_data/fish_hab.rds")
hab_dict = read_rds("../data/derived_data/hab_dict.rds")
qrf_data = read_rds("../data/derived_data/qrf_data.rds")
dens_offset = read_rds("../data/derived_data/dens_offset.rds")
qrf_mod = read_rds("../data/derived_data/qrf_mod.rds")
sel_hab_mets = read_rds("../data/derived_data/sel_hab_mets.rds")
gaa_covars = read_rds("../data/derived_data/gaa_covars.rds")
mod_data_weights = read_rds("../data/derived_data/mod_data_weights.rds")
all_preds = read_rds("../data/derived_data/all_preds.rds")
spawn_rec_params = read_rds("../data/derived_data/spawn_rec_params.rds")
model_svy_df = read_rds("../data/derived_data/model_svy_df.rds")
champ_wtsds = read_rds("../data/derived_data/champ_wtds.rds")
```

@@

To predict capacity at larger spatial scales, such as the watershed or population, we developed two extrapolation models (linear and areal capacity) based on globally available attributes (GAA) which were available for the entirety of tributary habitat utilized by a given population. The GAA data used here was taken from the list of GRTS master sample sites that the CHaMP sites were originally selected from [@Larsen2016]. Possible covariates included temperature range, growing degree days, an index of disturbance, the square root of cumulative drainage area, stream power, slope, channel type and watershed (Table \@ref(tab:gaa-covars)). The natural log of the CHaMP site capacity predictions was used as the response variable in a multiple linear regression model that incorporated the design weights of the CHaMP sites using the *svyglm* function from the *survey* package [@R-survey] in R software [@Rsoftware2019]. We fit two extrapolation models to each type of fish density, one that included watershed as covariate, for use in predicting capacity within CHaMP watersheds, and one that did not, for predicting everywhere else. We then made predictions of linear and areal capacity at all master sample sites throughout the interior CRB, generally spaced about one kilometer apart. Summaries of extrapolation model fit are shown in Table \@ref(tab:extrap-r2-table).

@@

<!-- Try to include images of high capacity / low capacity sites  -->
See Figures \@ref{fig:site-photo-high} and \@ref{fig:site-photo-low}.

@@

@@

```{r select-mets-table}
qrf_mod$importance %>%
  as_tibble(rownames = 'ShortName') %>%
  rename(imp = IncNodePurity) %>%
  arrange(desc(imp)) %>%
  mutate(Rank = 1:n()) %>%
  select(-imp) %>%
  left_join(hab_dict) %>%
  filter(MetricGroupName != 'Tier 1 Summary') %>%
  select(Rank,
         Metric = Name,
         `Metric Category` = MetricCategory,
         Description = DescriptiveText) %>%
  # pander(caption = "(\\#tab:select-mets-table) Habitat metrics and descriptions of metrics included in the QRF model to predict spring/summer Chinook salmon parr capacity (during summer months). Metrics are ranked in order of relative importance.")
  kable(caption = "Habitat metrics and descriptions of metrics included in the QRF model to predict spring/summer Chinook salmon parr capacity (during summer months). Metrics are ranked in order of relative importance.",
        booktabs = T,
        format.args = list(width = c(2, 10, 10, 50)))

@@

```{r rel-imp-figure, fig.cap = "Relative importance of each habitat covariate included in the quantile regression forest (QRF) model to predict habitat capacity, during summer months, for spring/summer Chinook salmon parr"}
qrf_mod$importance %>%
  as_tibble(rownames = 'ShortName') %>%
  rename(imp = IncNodePurity) %>%
  arrange(desc(imp)) %>%
  left_join(hab_dict %>%
              filter(is.na(MetricGroupName) | MetricGroupName != 'Tier 1 Summary')) %>%
  mutate_at(vars(ShortName, Name, MetricCategory),
            list(~ fct_reorder(., .x = imp))) %>%
  ggplot(aes(x = Name,
             y = imp)) +
  # geom_col(aes(fill = MetricCategory)) +
  # scale_fill_brewer(palette = 'Set2') +
  geom_col(fill = 'gray40') +
  coord_flip() +
  labs(x = 'Metric',
       y = 'Relative Importance')

@@

```{r pdp-figure, fig.cap = "Partial dependence plots for the spring/summer Chinook salmon parr capacity quantile regression forest (QRF) model, depicting how parr capacity shifts as each habitat metric changes, assuming all other habitat metrics remain at their mean values. Tick marks along the X-axis depict observed values."}
plot_partial_dependence(rf_mod = qrf_mod,
                        data = qrf_data,
                        data_dict = hab_dict,
                        type = 'quantile',
                        pred_quantile = 0.9,
                        log_transform = T,
                        log_offset = dens_offset,
                        scales = 'free') +
  labs(y = 'Capacity Prediction (per m)') +
  scale_color_grey(start = 0.5, end = 0.5) +
  theme(legend.position = 'none',
        strip.text = element_text(size = 6))
```

>>>>

```{r load-data}
fish_hab = read_rds("../data/derived_data/fish_hab.rds")
hab_dict = read_rds("../data/derived_data/hab_dict.rds")
qrf_data = read_rds("../data/derived_data/qrf_data.rds")
dens_offset = read_rds("../data/derived_data/dens_offset.rds")
qrf_mod = read_rds("../data/derived_data/qrf_mod.rds")
sel_hab_mets = read_rds("../data/derived_data/sel_hab_mets.rds")
gaa_covars = read_rds("../data/derived_data/gaa_covars.rds")
mod_data_weights = read_rds("../data/derived_data/mod_data_weights.rds")
all_preds = read_rds("../data/derived_data/all_preds.rds")
spawn_rec_params = read_rds("../data/derived_data/spawn_rec_params.rds")
model_svy_df = read_rds("../data/derived_data/model_svy_df.rds")
champ_wtsds = read_rds("../data/derived_data/champ_wtds.rds")

sel_hab_mets %<>%
  left_join(hab_dict %>%
              select(Metric = ShortName,
                     Name) %>%
              distinct()) %>%
  mutate(Name = recode(Name,
                       'Wetted Width To Depth Ratio Avg' = 'Width:Depth Ratio',
                       'Wetted Depth SD' = 'Depth Complexity',
                       'Wetted Channel Braidedness' = 'Braidedness',
                       'Fish Cover: Total' = 'Fish Cover',
                       'Riparian Cover: Some Canopy' = 'Riparian Canopy',
                       'Wetted Width Integrated' = 'Wetted Width',
                       'Substrate < 6mm' = 'Fines',
                       # 'Substrate: D16',
                       'Avg. August Temperature' = 'Avg. August Temp.',
                       'Large Wood Volume: Wetted Slow Water' = 'Large Wood Freq. in Pools'))

@@

To predict capacity at larger spatial scales, such as the watershed or population, we developed two extrapolation models (linear and areal capacity) based on globally available attributes (GAA) which were available for the entirety of tributary habitat utilized by a given population. The GAA data used here was taken from the list of GRTS master sample sites that the CHaMP sites were originally selected from [@Larsen2008; @Larsen2016]. Possible covariates included temperature range, growing degree days, an index of disturbance, the square root of cumulative drainage area, stream power, slope, channel type and watershed (Table \@ref(tab:gaa-covars)). The natural log of the CHaMP site capacity predictions was used as the response variable in a multiple linear regression model that incorporated the design weights of the CHaMP sites using the *svyglm* function from the *survey* package [@R-survey] in R software [@Rsoftware2019]. We fit two extrapolation models to each type of fish density, one that included watershed as covariate, for use in predicting capacity within CHaMP watersheds, and one that did not, for predicting everywhere else. We then made predictions of linear and areal capacity at all master sample sites throughout the interior CRB, generally spaced about one kilometer apart. Summaries of extrapolation model fit are shown in Table \@ref(tab:extrap-r2-table).

@@

<!-- Try to include images of high capacity / low capacity sites  -->
See Figures \@ref(fig:site-photo-high) and \@ref(fig:site-photo-low).

@@

<!-- The beginning of the following paragraph may need some word-smithing -->

@@

```{r select-mets-table}
qrf_mod$importance %>%
  as_tibble(rownames = 'ShortName') %>%
  rename(imp = IncNodePurity) %>%
  arrange(desc(imp)) %>%
  mutate(Rank = 1:n()) %>%
  select(-imp) %>%
  left_join(hab_dict %>%
              select(-Name) %>%
              inner_join(sel_hab_mets %>%
                           rename(ShortName = Metric))) %>%
  filter(MetricGroupName != 'Tier 1 Summary') %>%
  select(Rank,
         Metric = Name,
         `Metric Category` = MetricCategory,
         Description = DescriptiveText) %>%
  # pander(caption = "(\\#tab:select-mets-table) Habitat metrics and descriptions of metrics included in the QRF model to predict spring/summer Chinook salmon parr capacity (during summer months). Metrics are ranked in order of relative importance.")
  kable(caption = "Habitat metrics and descriptions of metrics included in the QRF model to predict spring/summer Chinook salmon parr capacity (during summer months). Metrics are ranked in order of relative importance.",
        booktabs = T,
        format.args = list(width = c(2, 10, 10, 50)))

@@

```{r rel-imp-figure, fig.cap = "Relative importance of each habitat covariate included in the quantile regression forest (QRF) model to predict habitat capacity, during summer months, for spring/summer Chinook salmon parr"}
qrf_mod$importance %>%
  as_tibble(rownames = 'ShortName') %>%
  rename(imp = IncNodePurity) %>%
  mutate(imp = (imp / max(imp)) * 100) %>%
  arrange(desc(imp)) %>%
  left_join(hab_dict %>%
              filter(is.na(MetricGroupName) | MetricGroupName != 'Tier 1 Summary')) %>%
  select(-Name) %>%
  left_join(sel_hab_mets %>%
              select(ShortName = Metric,
                     Name)) %>%
  mutate_at(vars(ShortName, Name, MetricCategory),
            list(~ fct_reorder(., .x = imp))) %>%
  ggplot(aes(x = Name,
             y = imp)) +
  # geom_col(aes(fill = MetricCategory)) +
  # scale_fill_brewer(palette = 'Set2') +
  geom_col(fill = 'gray40') +
  coord_flip() +
  labs(x = 'Metric',
       y = 'Relative Importance')

@@

```{r pdp-figure, fig.cap = "Partial dependence plots for the spring/summer Chinook salmon parr capacity quantile regression forest (QRF) model, depicting how parr capacity shifts as each habitat metric changes, assuming all other habitat metrics remain at their mean values. Tick marks along the X-axis depict observed values."}
plot_partial_dependence(rf_mod = qrf_mod,
                        data = qrf_data,
                        data_dict = hab_dict %>%
                          select(-Name) %>%
                          inner_join(sel_hab_mets %>%
                                       select(ShortName = Metric,
                                              Name)),
                        type = 'quantile',
                        pred_quantile = 0.9,
                        log_transform = T,
                        log_offset = dens_offset,
                        scales = 'free') +
  labs(y = 'Capacity Prediction (per m)') +
  scale_color_grey(start = 0.5, end = 0.5) +
  theme(legend.position = 'none',
        strip.text = element_text(size = 6))
```
